\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-04-08 bmc.tex
prelims changes

    \item \emph{S3CAM Fails}: No abstract counter-example is found by
        S3CAM. We can remedy it by increasing search budgets and/or
        restarting.

    \item \emph{BMC Fails}: An abstract counter-example is found, but
        the BMC fails to find a concrete counter-example in the PWA
        relational model. The failure can be attributed to either (a)
        a spurious abstract counter-examples or (b) a poorly estimated
        model. The former can be addressed by restarting but the
        latter requires that the maximum model error $\delta_{max}$ be
        decreased.

    \item \emph{Inaccurate PWA Modeling}: An abstract counter-example
        is found, and it is successfully concretized in the PWA
        relational model by the BMC. However, it is not reproducible
        in the black box system. This happens when the PWA relational
        model is not precise enough.
\end{itemize}

>>>>

    \item \emph{S3CAM Fails} No abstract counter-example is found by
        S3CAM. We can remedy it by increasing search budgets and/or
        restarting.

    \item \emph{BMC Fails} An abstract counter-example is found, but
        the BMC fails to find a concrete counter-example in the PWA
        relational model. The failure can be attributed to either (a)
        a spurious abstract counter-examples or (b) a poorly estimated
        model. The former can be addressed by restarting but the
        latter requires that the maximum model error $\delta_{max}$ be
        decreased.

    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
        is found, and it is successfully concretized in the PWA
        relational model by the BMC. However, it is not reproducible
        in the black box system. This happens when the PWA relational
        model is not precise enough.

\end{itemize}

### 2017-04-08 intro.tex
prelims changes

In many settings, we may not have a physics-based model derived from
first principles, but may just have time-series data of the system
behavior. In such a setting, a data-driven model allows us to
generalize individual time-series behaviors. But, there are several
approaches in the literature to perform data-driven modeling; e.g.,
system identification techniques that learn dynamic models from data
\cite{ljung1999system}, auto-regressive models for forecasting
\cite{wei1994time}, machine learning techniques that learn static or
dynamic models from data
\cite{narendra1990identification,lu2009linear}.  Why then do we need a
new modeling technique? The answer to this question lies in our goal
for the second step, \ie, symbolic analysis of the learned model. For
this step, we need to have models that can be digested by existing
symbolic tools, which is harder to do with some of the aforementioned
heavier data-driven modeling methods. Hence, we use a lighter flavor
of data-driven modeling such as the one proposed for learning
relational models from data
\cite{zutshi2012timed,sankaranarayanan2011relational}. By learning a
{\em PWA relational model}, \ie, essentially a PWA discrete-time
dynamical system, we can use software analysis approaches such as
bounded model checking assisted by SMT solvers and linear programming
solvers.

@@

In \cite{zutshi2014multiple}, the authors investigate an on-the-fly
abstraction technique that is close to the approach we propose. A key
difference is that the abstraction explored in this work is an
existential {\em relational} abstraction. The approximate model
learned is a directed acyclic graph, where nodes represent regions of
the state-space, and an edge between two regions indicates a
transition between states in the regions. Our approach further
improves the existential abstraction by annotating the edges in the
graph-based abstraction with affine relations between the source and
destination states.

% We further our exploration of trajectory segment based methods; and
% explore an alternative approach to overcome the explosion in abstract
% states. Instead of selectively refining the abstraction, we compute a
% model of the black box system and use bounded model checking to find a
% concrete counter-example in the model. Due to modeling errors, this
% might not be reproducible in the original black-box system.

More specifically, we use linear regression to quantitatively estimate
the discovered edges in the graph-based abstraction of the
state-space. These linear maps approximate locally observed behaviors,
and the resulting graph or the Piece-Wise Affine (PWA) relational
model can be interpreted as an infinite state discrete transition
system and model checked for time bounded safety properties. A
counter-example if found, can indicate the presence of a violation in
the system.  We then use this counter-example to guide the search
towards a counter-example in the original system.

>>>>

In many settings, we may not have a physics-based model derived from
first principles, but may just have time-series data of the system
behavior. In such a setting, a data-driven model allows us to
generalize individual time-series behaviors. But, there are several
approaches in the literature to perform data-driven modeling; e.g.,
system identification techniques that learn dynamic models from data
\cite{ljung1999system}, auto-regressive models for forecasting
\cite{wei1994time}, bounded-error hybrid systems identification
methods \cite{bemporad2003greedy,bemporad2005bounded}, clustering
based methods \cite{ferrari2003clustering}, methods to identify
piecewise affine dynamics \cite{paoletti2007identification}, and
machine learning techniques that learn static or dynamic models from
data
\cite{narendra1990identification,lu2009linear,juloski2005bayesian}.
Why then do we need a new modeling technique? The answer to this
question lies in our goal for the second step, \ie, symbolic analysis
of the learned model. For this step, we need to have models that can
be digested by existing symbolic tools, which is harder to do with
some of the aforementioned heavier data-driven modeling methods.
Hence, we use a lighter flavor of data-driven modeling such as the one
proposed for learning relational models from data
\cite{zutshi2012timed,sankaranarayanan2011relational}. By learning a
{\em PWA relational model}, \ie, essentially a PWA discrete-time
dynamical system, we can use software analysis approaches such as
bounded model checking assisted by SMT solvers and linear programming
solvers.

@@

In \cite{zutshi2014multiple}, the authors investigate an on-the-fly
abstraction technique that is close to the approach we propose.  Both
approaches use an existential abstraction of the state-space of the
hybrid dynamical system. The abstract model learned is a directed
acyclic graph, where nodes represent regions of the state-space, and
an edge between two regions indicates a transition between states in
the regions.  A key difference in this work is that the model being
learned in addition to being an existential graph-based abstraction is
also a {\em relational} abstraction.  In other words, the approach
presented here improves the existential abstraction by annotating the
edges in the graph-based abstraction with affine relations between the
source and destination states.  More specifically, we use linear
regression to quantitatively estimate the discovered edges in the
graph-based abstraction of the state-space. These linear maps
approximate locally observed behaviors, and the resulting graph or the
Piece-Wise Affine (PWA) relational model can be interpreted as an
infinite state discrete transition system and model checked for time
bounded safety properties. A counter-example if found, can indicate
the presence of a violation in the system.  We then use this
counter-example to guide the search towards a counter-example in the
original system.

We remark that our work focusses on building discrete PWA models, and
is unrelated to work on analyzing dynamical systems that have PWA
vector fields. Techniques to analyze PWA dynamical systems using
bisimulation relations translate the system into a non-deterministic
infinite state transition system, and use model checking approaches to
verify them \cite{asarin2000approximate}. Related approaches using
bisimulation relations \cite{pappas2003bisimilar,tabuada2006linear,
yordanov2007model} have interesting parallels with the work presented
here.

### 2017-04-08 prelims.tex
emsoft formatvi main.tex
.
prelims changes

\subsection{Piecewise Affine (PWA) Transition System}

We define the PWA transition system as a discrete transition system
$\rho:\tupleof{\pLocs,\vx,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
a nonempty set of locations, $\vx$ is a set of variables, $\scrT$
is a finite set of transitions of the form $\tupleof{l, \tau,
l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
post-location and $\tau \in \scrT$ is the transition. $\tau$
is labled by the transition relation $\rho_{\tau}$
over the current states and the next states. The transition relation
is defined in terms of affine predicates on the current and next
states and an affine update between them.
\[
    \rho_{\tau}(\vx,\vx') \subseteq \setof{(\vx,\vx')) \;|\;
    g_\tau(\vx) \land g'_\tau(\vx') \land f_{\tau}(\vx, \vx')}
\]
where $f_{\tau}$ is an affine relation on the pre and post states
$\vx,\vx'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
post conditions) on the states. Finally, $\pLocI$ is an initial
location $\Theta$ is an affine assertion characterizing the initial
states included in $\pLocI$.

@@

We can now use a PWA transition system $\rho$ to approximate the
behavior of a hybrid dynamical system $\System$ defined by a simulator
$\simulate^{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
form, we can only approximate general hybrid systems.}. Each
transition of $\rho$ represents a discrete time step $\Delta$. Let the
state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
predicates $g_\tau$ are given by a conjunction of hyper-planes and
hence can be represented in matrix form as a polyhedron (defined by
$m$ constraints) in the state-space
\[
g:C\x - \vd \le 0
\]
where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
$m$. The affine relation $f_\tau$ can be represented using matrix $A$
and an offset vector $\vb$ as $f_\tau \subseteq \setof{(\x,\x') | \x'
= A\x+\vb}$. Finaly, we incorporte an approximate non-deterministic
error term $\delta$, defined as a vector of intervals
$[\delta_{\min},\delta_{\max}]$.
\[
    f_\tau \subseteq \setof{(\x,\x') | \x' = A\x+\vb}.
\]
%defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
%\in f_\tau(\x)}$.

@@

Given a hybrid dynamical system over a state-space $\reals^n$, a PWA
relational model $\rho$ is given by the tuple
$\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in \scrT$ is a
collection of affine transitions  and $\Theta$ is an affine predicate
over $\x\in\real^n$. The transition relation is then defined by
$\scrT$ with $n$ transition relations as follows

\begin{equation}
    \scrT = \left\{
        \begin{array}{ll}
            g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
            \ldots\\
            g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
        \end{array}
    \right.
\end{equation}
\end{definition}

where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
relations and $\x'$ denotes the next state of the system.

A PWA model is \textit{deterministic}, iff for every state
$\x\in\reals^n$, a unique guarded affine map is satisfied and if all
entries in the error interval vector $\vec{\delta}$ are singletons.
However, in practice such a case rarely exists. A PWA model will be
usually non-deterministic, both due to multiple choices of transitions
from any location and the reachable states at the $k^{th}$ time step
(or after $k\Delta$ units of time) being a set.  Abusing the notation,
we denote the set of states reachable in a single step in $\rho$ by
$\x' = \rho(\x)$.

A PWA transition system is \textit{complete}, iff for every state
$\x\in\HybridStates^n$, there exists at least one satisfied
transition relation $\scrT$. If this is not the case, the system can
deadlock, with no further executions. This usually results from
modeling errors, and from here on, we do not consider such
cases.

>>>>

\subsection{PWA Transition System}

We define a PWA transition system $A$ as a tuple
$\tupleof{\pLocs,\vx,\scrT, G, U, \pLocI,\Theta}$, where:
\begin{itemize}[noitemsep, leftmargin= 1.5 em]
\item
$\pLocs$ is a set of locations,
\item
$\vx$ is a variable that takes values in $\reals^m$,
\item
$\scrT$ is a finite subset of $\pLocs \times G \times U \times G \times \pLocs$, 
\item
$G$ is a set of conjunctions of affine predicates, 
\item
$U$ is a set of affine relations,
\item
$\pLocI$ is an initial location, and,
\item
$\Theta$ is an initial affine predicate.
\end{itemize}

The operation of $A$ is described in terms of its configuration and
its moves. A configuration of $A$ is a pair $(\ell, \nu)$, where $\ell
\in \pLocs$, and $\nu$ is a valuation function assigning values in
$\reals^m$ to the variable $\vx$.  The move function $\mu$ maps a
configuration $(\ell,\nu)$ to the next configuration $(\ell',\nu')$,
iff there is a transition $\tau$ of the form $(\ell,g,f,g',\ell')$,
and $\nu(\vx)$ satisfies predicate $g$, $\nu'(\vx)$ satisfies
predicate $g'$ and $\nu'(\vx) \in f(\nu(\vx))$.

For a given $\tau$, we call the affine relation $f_\tau$ its update
relation, and the predicates $g_\tau$ and $g'_\tau$ as pre- and
post-conditions on the transition. The machine starts its operation in
location $\pLocI$ with an initial valuation $\nu_0$ to the variable
$\vx$ such that $\nu_0(\vx)$ satisfies $\Theta$.

% We define the PWA transition system as a discrete transition system
% $\rho:\tupleof{\pLocs,\vx,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
% a nonempty set of locations, $\vx$ is a set of variables, $\scrT$
% is a finite set of transitions of the form $\tupleof{l, \tau,
% l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
% post-location and $\tau \in \scrT$ is the transition. $\tau$
% is labled by the transition relation $\rho_{\tau}$
% over the current states and the next states. The transition relation
% is defined in terms of affine predicates on the current and next
% states and an affine update between them.
% \[
%     \rho_{\tau}(\vx,\vx') \subseteq \setof{(\vx,\vx')) \;|\;
%     g_\tau(\vx) \land g'_\tau(\vx') \land f_{\tau}(\vx, \vx')}
% \]
% where $f_{\tau}$ is an affine relation on the pre and post states
% $\vx,\vx'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
% post conditions) on the states. Finally, $\pLocI$ is an initial
% location $\Theta$ is an affine assertion characterizing the initial
% states included in $\pLocI$.

@@

We how we can use the formal machinery of a PWA transition system $A$
to approximate the behavior of a hybrid dynamical system $\System$
defined by a simulator $\simulate^{\System}$~\footnote{Due to $f_\tau$
being restricted to an affine form, we can only approximate general
hybrid systems.}.

%defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
%\in f_\tau(\x)}$.

@@

A PWA relational model $A(\System)$ of a hybrid dynamical system
$\System$ is simply a PWA transition system, where:

\begin{itemize}[noitemsep, leftmargin= 1.5 em]
\item
each transition of $A$ represents a discrete time step $\Delta$ of the
system $\System$,
\item
if the state-space of $\System$ is $\reals^n$, the the valuation
function for variable $\x$ in $A$ assigns $\x$ some value in
$\reals^n$,
\item
for the $N$ tuples $(v,v')$ discovered by simulation, if $v' = Av
+ B + \vec{\delta}$ represents the affine relation discovered using
linear regression, and if $h$ is the convex hull of all the $v$
states across all tuples and $h'$ is the convex hull of all the $v'$
states across all tuples, then $f_\tau \subseteq
\setof{(\nu(\x),\nu'(\x)) \mid \nu'(\x) \in A\x + B + \vec{\delta}}$,
and $g$ states that $\nu(\x)$ is in $h$, and $g'$ states that
$\nu'(\x) \in h'$.
\end{itemize}

Note that the guard predicates $g_\tau$ represent a conjunction of $m$
affine inequalities, then we can represent $g_\tau$ as follows:
\[
g_\tau  \equiv C\x - \vd \le 0
\]
where $C$ is an $m \times n$ matrix and $\vd$ is $m \times 1$ matrix.

% Given a hybrid dynamical system over a state-space $\reals^n$, a PWA
% relational model $\rho$ is given by the tuple
% $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in \scrT$ is a
% collection of affine transitions  and $\Theta$ is an affine predicate
% over $\x\in\real^n$. The transition relation is then defined by
% $\scrT$ with $n$ transition relations as follows
% 
% \begin{equation}
%     \scrT = \left\{
%         \begin{array}{ll}
%             g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
%             \ldots\\
%             g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
%         \end{array}
%     \right.
% \end{equation}
\end{definition}

% where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
% relations and $\x'$ denotes the next state of the system.

A PWA relational model is \textit{deterministic}, iff for every state
$\x\in\reals^n$, a unique guarded affine map is satisfied and if all
entries in the error interval vector $\vec{\delta}$ are singletons.
However, in practice such a case rarely exists. A PWA model will be
usually non-deterministic, both due to multiple choices of transitions
from any location and the reachable states at the $k^{th}$ time step
(or after $k\Delta$ units of time) being a set.  Abusing the notation,
we denote the set of states reachable in a single step in $\rho$ by
$\x' = \rho(\x)$.

A PWA relational model system is \textit{complete}, iff for every
state $\x\in\HybridStates^n$, there exists at least one satisfied
transition relation $\scrT$. If this is not the case, the system can
deadlock, with no further executions. This usually results from
modeling errors, and from here on, we do not consider such cases.
